---
title: "ICE I-213 analysis 2022-2025"
author: "UWCHR"
date: "2026-02-18"
output:
    html_document:
        html_preview: true
        toc: true
        toc_depth: 3
        toc_float: true
        code_folding: hide
---

```{r setup, message=FALSE, warning=FALSE, include=TRUE}

options(scipen = 1000000)
options(ggplot2.continuous.colour = "viridis")
options(ggplot2.continuous.fill = "viridis")
options(ggplot2.discrete.colour = "viridis")
options(ggplot2.discrete.fill = "viridis")

library(pacman)

p_load(here, tidyverse, zoo, scales, lubridate, ggplot2, plotly, gghighlight, viridis, ggmap, tidygeocoder, sf, rnaturalearth)

every_nth = function(n) {
  return(function(x) {x[c(TRUE, rep(FALSE, n - 1))]})
}

every_2nd_nth = function(n) {
  return(function(x) {x[c(FALSE, rep(TRUE, n - 1))]})
}

```


```{r load_data, message=FALSE, warning=FALSE, include=TRUE}

file <- "sea-i213s-2026-02-13-geocoded.csv"

df <- read_delim(here('analyze', 'input', file), delim='|')

ddp_arrests <- arrow::read_feather(
  "https://github.com/deportationdata/ice/raw/refs/heads/main/data/arrests-latest.feather"
)

ddp_arrests_sea <- ddp_arrests %>% 
  mutate(apprehension_date = as.Date(apprehension_date),
         final_order_date = as.Date(final_order_date)) |>
  mutate(apprehension_method = fct(apprehension_method)) |>
  mutate(criminality = apprehension_criminality == "1 Convicted Criminal") |>
  mutate(n_dupe = row_number(), .by = c("apprehension_date", "apprehension_method", "apprehension_criminality", "unique_identifier")) |>
  filter(is.na(unique_identifier) | n_dupe == 1) |>
  mutate(week = floor_date(apprehension_date, "week", week_start = "Monday"),
         month = floor_date(apprehension_date, "month"),
         cy_quarter = as.factor(quarter(apprehension_date, fiscal_start = 1, with_year=TRUE)),
         fy_quarter = as.factor(quarter(apprehension_date, fiscal_start = 10, with_year=TRUE)),
         fy = substr(quarter(apprehension_date, fiscal_start = 10, with_year = TRUE), 0, 4),,
         cy = year(apprehension_date)
         ) %>% 
  filter(apprehension_aor == "Seattle Area of Responsibility")

# rm(ddp_arrests)

```



```{r}

dat1 <- df %>% 
  filter(area_of_responsibility_code == "SEA",
         month >= "2022-01-01") %>% 
  count(month) %>% 
  mutate(source = "SEA I-213s")

dat2 <- ddp_arrests_sea %>% 
  filter(apprehension_aor == "Seattle Area of Responsibility",
         month >= "2022-01-01") %>% 
  count(month) %>% 
  mutate(source = "SEA arrests per DDP")

dat <- rbind(dat1, dat2)

p1 <- dat %>% 
  ggplot(aes(x = month, y = n, color = source)) +
  geom_line() +
  labs(title = "I-213s versus ICE arrests")

p1

# Comparison of DDP arrests to I-213s per month for SEA AOR

dat_wide <- pivot_wider(dat, id_cols = month, names_from = source, values_from = n) %>% 
  janitor::clean_names() %>% 
  drop_na() %>% 
  filter(month < "2025-10-01") %>% 
  mutate(diff = sea_arrests_per_ddp - sea_i_213s,
         pct = sea_i_213s / sea_arrests_per_ddp * 100)

```

```{r}

# uwchr <- read_delim(here::here('analyze', 'input', 'ice_arrests_fy12-23ytd.csv.gz'), delim='|')

url <- "https://drive.usercontent.google.com/download?id=1XAOS7t3ijImVXQ9G7l6xdCr_rBh8i4cz&export=download&authuser=0&confirm=t&uuid=0c4a4b74-5e5f-4f00-8c8a-eb00a56708f9&at=APcXIO3JvUiU3ilszJZ3EAuIE01Y:1772065031530"
f <- tempfile(fileext = ".csv.gz")
download.file(url, f, mode = "wb")

uwchr <- read.delim(f, sep='|')

uwchr_sea <- uwchr %>% 
  filter(area_of_responsibility == "Seattle Area of Responsibility") %>%
  mutate(arrest_date = mdy(arrest_date),
         month = floor_date(arrest_date, unit="month"),
         cy_quarter = as.factor(quarter(arrest_date, fiscal_start = 1, with_year = TRUE)),
         fy_quarter = as.factor(quarter(arrest_date, fiscal_start = 10, with_year = TRUE)),
         fy = as.numeric(substr(quarter(arrest_date, fiscal_start = 10, with_year = TRUE), 0, 4)),
         cy = year(arrest_date),
         source = "All arrests (UWCHR)")

rm(uwchr)

```

```{r}
  
uwchr_sea_per_fy_quarter <- uwchr_sea %>% 
  count(fy_quarter, source) %>% 
  filter(fy_quarter != "2023.2")
  
i213s_per_fy_quarter <- df %>% 
  mutate(source = "I-213s (UWCHR)") %>% 
   filter(month >= "2022-01-01") %>%
  count(fy_quarter, source)

ddp_sea_per_fy_quarter <- ddp_arrests_sea %>% 
  filter(apprehension_aor == "Seattle Area of Responsibility") %>% 
  count(fy_quarter) %>% 
  filter(fy_quarter != "2023.4",
         fy_quarter != "2026.1") %>% 
  mutate(source = "All arrests (DDP)")

dat <- rbind(i213s_per_fy_quarter, ddp_sea_per_fy_quarter, uwchr_sea_per_fy_quarter) %>% 
  mutate(fy_quarter = as.factor(as.character(fy_quarter)))

every_nth = function(n) {
  return(function(x) {x[c(TRUE, rep(FALSE, n - 1))]})
}

p1 <- dat %>% 
  mutate(fy_quarter = str_replace_all(fy_quarter, "\\.", "Q")) %>% 
  ggplot(aes(x = fy_quarter, y = n, fill = source)) +
  geom_col(position = position_dodge2(preserve = "single")) +
  scale_x_discrete(breaks = every_nth(n = 8)) +
  labs(title = "Historic increase in ICE apprehensions in\nQ4 of CY 2026",
       subtitle = "Seattle Area of Responsibility")

p1

p2 <- dat %>% 
  mutate(fy_quarter = str_replace_all(fy_quarter, "\\.", "Q"),
         source = factor(source, levels = c("I-213s (UWCHR)", "All arrests (DDP)", "All arrests (UWCHR)"))
         ) %>% 
  ggplot(aes(x = fy_quarter, y = n, linetype =source, group = source),color = viridis(1)) +
  geom_line() +
  ylab("") +
  ylim(0, 2500) +
  xlab("Fiscal quarter") +
  scale_x_discrete(breaks = every_nth(n = 8)) +
  labs(title = "ICE apprehensions in late 2025 approach historical high point",
       subtitle = "Seattle Area of Responsibility, Oct. 2011 - Dec. 2025",
       linetype = "Data source",
       caption = "Sources: I-213 dataset obtained by UWCHR,\nICE arrests datasets obtained by UWCHR, Deportation Data Project.") +
  theme_minimal()

p2


```

```{r}
  
uwchr_sea_per_cy_quarter <- uwchr_sea %>% 
  mutate(cy_quarter_zoo = as.yearqtr(arrest_date, format = "%Y-%M-%D")) %>% 
  count(cy_quarter_zoo, source) %>% 
  filter(cy_quarter_zoo != "2023-1")
  
i213s_per_cy_quarter <- df %>% 
  mutate(cy_quarter_zoo = as.yearqtr(apprehension_date, format = "%Y-%M-%D"),
    source = "I-213s (UWCHR)") %>% 
   filter(apprehension_date >= "2022-01-01") %>%
  count(cy_quarter_zoo, source)

ddp_sea_per_cy_quarter <- ddp_arrests_sea %>% 
  mutate(cy_quarter_zoo = as.yearqtr(apprehension_date, format = "%Y-%M-%D")) %>% 
  filter(apprehension_aor == "Seattle Area of Responsibility",) %>% 
  count(cy_quarter_zoo) %>% 
  filter(cy_quarter_zoo != "2023-3",
         cy_quarter_zoo != "2025-4") %>% 
  mutate(source = "All arrests (DDP)")

dat <- rbind(i213s_per_cy_quarter, ddp_sea_per_cy_quarter, uwchr_sea_per_cy_quarter)

p1 <- dat %>% 
  mutate(source = factor(source, levels = c("I-213s (UWCHR)", "All arrests (DDP)", "All arrests (UWCHR)"))
         ) %>% 
  ggplot(aes(x = cy_quarter_zoo, y = n, linetype = source, group = source)) +
  geom_line(color = viridis(1)) +
  ylab("Arrests") +
  ylim(0, 2500) +
  xlab("") +
  scale_x_yearqtr(format = "%Y") +
  labs(title = "ICE apprehensions in late 2025 approach historical high point",
       subtitle = "Seattle Area of Responsibility, Oct. 2011 - Dec. 2025",
       linetype = "Data source",
       caption = "Sources: I-213 dataset obtained by UWCHR,\nICE arrests datasets obtained by UWCHR, Deportation Data Project.") +
  theme_minimal()

p1


```

# Comparison of counts per country of citizenship

I-213 records are produced at the time of initial apprehension of someone deemed "deportable/inadmissible"; they do not appear to reflect later arrests of people released on parole, electronic detention, or other forms of alternatives to detention. In the below, we analyze the differences in counts of arrests by country of citizenship for the Seattle AOR during the time period of overlap between the I-213 data and data obtained by the Deportation Data Project. Discrepancies in total arrest counts vary widely by nationality in ways that may map onto historic DHS policy shifts and may obscure actual trends in contemporary arrests.

```{r}

ddp_country_per_cy_q <- ddp_arrests_sea %>% 
  filter(apprehension_date >= "2023-10-01",
         apprehension_date <= "2025-07-15") %>% 
  count(cy_quarter, citizenship_country) %>% 
  mutate(source = "DDP")

i213_country_per_cy_q <- df %>% 
  filter(apprehension_date >= "2023-10-01",
       apprehension_date <= "2025-07-15") %>% 
  count(cy_quarter, citizenship_country_descr) %>% 
  rename(citizenship_country = citizenship_country_descr) %>% 
  mutate(source = "i213s")

dat <- rbind(ddp_country_per_cy_q, i213_country_per_cy_q)

dat_wide <- dat %>% pivot_wider(id_cols = c('cy_quarter', 'citizenship_country'), names_from = 'source', values_from = 'n', values_fill = 0) %>% 
  mutate(diff = DDP - i213s,
         pct_missing = diff / DDP)

```

```{r}

ddp_country_per_cy <- ddp_arrests_sea %>% 
  filter(apprehension_date >= "2023-10-01",
         apprehension_date <= "2025-07-15") %>% 
  count(cy, citizenship_country) %>% 
  mutate(source = "DDP")

i213_country_per_cy <- df %>% 
  filter(apprehension_date >= "2023-10-01",
       apprehension_date <= "2025-07-15") %>% 
  count(cy, citizenship_country_descr) %>% 
  rename(citizenship_country = citizenship_country_descr) %>% 
  mutate(source = "i213s")

dat <- rbind(ddp_country_per_cy, i213_country_per_cy)

dat_wide <- dat %>%
  filter(n > 10) %>% 
  pivot_wider(id_cols = c('cy', 'citizenship_country'),
              names_from = 'source',
              values_from = 'n',
              values_fill = 0) %>% 
  mutate(diff = DDP - i213s,
         pct_missing = diff / DDP)

```

```{r}

ddp_country <- ddp_arrests_sea %>% 
  filter(apprehension_date >= "2023-10-01",
         apprehension_date <= "2025-07-15") %>% 
  count(citizenship_country) %>% 
  mutate(source = "DDP")

i213_country <- df %>% 
  filter(apprehension_date >= "2023-10-01",
       apprehension_date <= "2025-07-15") %>% 
  count(citizenship_country_descr) %>% 
  rename(citizenship_country = citizenship_country_descr) %>% 
  mutate(source = "i213s")

dat <- rbind(ddp_country, i213_country)

dat_wide <- dat %>%
  pivot_wider(id_cols = c('citizenship_country'),
              names_from = 'source',
              values_from = 'n',
              values_fill = 0) %>% 
  mutate(diff = DDP - i213s,
         pct_missing = diff / DDP)

```


```{r}

vdigest <- Vectorize(digest::digest)

ddp_arrests_sea <- ddp_arrests_sea %>% 
  mutate(gender_code = substr(gender, 0, 1),
         apprehension_site_landmark = str_trim(toupper(apprehension_site_landmark)),
         apprehension_method = str_trim(toupper(apprehension_method)),
         citizenship_country = str_trim(toupper(citizenship_country)))

df <- df %>% 
  filter(apprehension_date >= min(ddp_arrests$apprehension_date) &
           apprehension_date <= max(ddp_arrests$apprehension_date)) %>% 
    mutate(apprehension_landmark_descr = str_trim(toupper(apprehension_landmark_descr)),
         apprehension_method_descr = str_trim(toupper(apprehension_method_descr)),
         citizenship_country_descr = str_trim(toupper(citizenship_country_descr)),
         birth_year = year(apprehension_date) - age)

df <- df %>% rowwise() %>%
  unite(allCols, sep = "", remove = FALSE) %>%
  mutate(hash = vdigest(allCols)) %>%
  select(-c(allCols)) %>%
  arrange(desc(hash))

ddp_arrests_sea <- ddp_arrests_sea %>% rowwise() %>%
  unite(allCols, sep = "", remove = FALSE) %>%
  mutate(hash = vdigest(allCols)) %>%
  select(-c(allCols)) %>%
  arrange(desc(hash))

```

```{r}
i213_cols <- c("citizenship_country_descr",
               "gender_code",
               "apprehension_date",
               "apprehension_method_descr",
               "apprehension_landmark_descr",
               "birth_year"
               )

df <- df %>% rowwise() %>%
  unite(i213_cols, all_of(i213_cols), sep = "", remove = FALSE) %>%
  mutate(join_hash = vdigest(i213_cols)) %>%
  select(-c(i213_cols))

i_213_ambig <- df[duplicated(df$join_hash) | duplicated(df$join_hash, fromLast = TRUE), ] %>% 
  arrange(desc(join_hash))

sum(duplicated(df$join_hash))

ddp_cols <- c("citizenship_country",
              "gender_code",
              "apprehension_date",
              "apprehension_method",
              "apprehension_site_landmark",
              "birth_year"
              )

ddp_arrests_sea <- ddp_arrests_sea %>% rowwise() %>%
  unite(ddp_cols, all_of(ddp_cols), sep = "", remove = FALSE) %>%
  mutate(join_hash = vdigest(ddp_cols)) %>%
  select(-c(ddp_cols))

sum(duplicated(ddp_arrests_sea[ddp_arrests_sea$apprehension_aor == "Seattle Area of Responsibility", "join_hash"]))

ddp_sea_ambig <-  ddp_arrests_sea[duplicated(ddp_arrests_sea$join_hash) | duplicated(ddp_arrests_sea$join_hash, fromLast = TRUE), ]

ddp_sea_ambig <- ddp_sea_ambig %>% 
  filter(apprehension_aor == "Seattle Area of Responsibility") %>% 
  arrange(desc(join_hash))

```

```{r}

test_join <- left_join(df, ddp_arrests, by = c("citizenship_country_descr" = "citizenship_country", "gender_code" = "gender_code", "apprehension_date" = "apprehension_date", "apprehension_method_descr" = "apprehension_method", "apprehension_landmark_descr" = "apprehension_site_landmark" ))

# This is not accurate because duplicates but most rows have at least 1 match
sum(!is.na(test_join$apprehension_aor)) / nrow(test_join)

test_join_distinct_id <- test_join %>% 
  filter(!is.na(unique_identifier)) %>% 
  distinct(unique_identifier, .keep_all=TRUE)
  

```