---
title: "ICE I-213 analysis"
author: "UWCHR"
date: "2025-09-22"
output:
    html_document:
        html_preview: true
        toc: true
        toc_depth: 3
        toc_float: true
        code_folding: hide
---

```{r setup, message=FALSE, warning=FALSE, include=TRUE}

options(scipen = 1000000)
options(ggplot2.continuous.colour = "viridis")
options(ggplot2.continuous.fill = "viridis")
options(ggplot2.discrete.colour = "viridis")
options(ggplot2.discrete.fill = "viridis")

library(pacman)

p_load(here, tidyverse, zoo, lubridate, ggplot2, plotly, gghighlight, viridis, readxl, ggmap, ggrepel, tidygeocoder, sf, rnaturalearth)

general_caption = "Source: DHS I-213 forms obtained via FOIA\nCounty population estimates via U.S. Census 2020-2024\n Analysis: UW Center for Human Rights"

```

I-213 forms document a subset of all ICE arrests, and may systematically underrepresent some populations. We recommend caution in interpreting this data without comparison to other sources. Gecoded arrest locations are approximate, they do not represent exact coordinates of the enforcement activity.

```{r load_data, message=FALSE, warning=FALSE, include=TRUE}

file <- "sea-i213s-2026-02-13-geocoded.csv"

df <- read_delim(here('analyze', 'input', file), delim='|')

annual_arrests_pc <- read_delim(here::here("per-capita", "output", "annual-arrests-per-capita-wa-or-2022-2025.csv"), delim = ",")

quarterly_arrests_pc <- read_delim(here::here("per-capita", "output", "quarterly-arrests-per-capita-wa-or-2022-2025.csv"), delim = ",")

field_offices <- read_delim(here::here("analyze", "hand", "pnw-field-offices.csv"), delim = "|")

field_offices <- st_as_sf(field_offices, coords = c("long", "lat"),
                   crs = 4326, agr = "constant")

wa_coords <- coord_sf(xlim = c(-125, -116.5), ylim = c(45, 49.5), expand = FALSE, crs = 4326)
or_coords <- coord_sf(xlim = c(-125, -116), ylim = c(41.75, 46.5), expand = FALSE, crs = 4326)
pnw_coords <- coord_sf(xlim = c(-125, -116), ylim = c(41.5, 49.5), expand = FALSE, crs = 4326)
pdx_coords <- coord_sf(xlim = c(-124.5, -120), ylim = c(44.5, 47), expand = FALSE, crs = 4326)

world <- ne_countries(scale = "medium", returnclass = "sf")

states <- st_as_sf(maps::map("state", crs = 4326, plot = FALSE, fill = TRUE))

states <- cbind(states, st_coordinates(st_point_on_surface(states)))
states$ID <- tools::toTitleCase(as.character(states$ID))

shared_pc_scale <- max(annual_arrests_pc$arrests_pc)
wa_pc_scale <- max(annual_arrests_pc[annual_arrests_pc$state == "WA", "arrests_pc"])
or_pc_scale <- max(annual_arrests_pc[annual_arrests_pc$state == "OR", "arrests_pc"])

```

```{r pnw_ice_field_offices, echo=TRUE, warning=FALSE}

pnw <- states %>% 
  filter(ID == c("Washington", "Oregon"))

counties <- st_as_sf(maps::map("county", plot = FALSE, fill = TRUE))

counties <- counties %>% 
  separate(col = 'ID', into = c("state", "county"), sep = ",", remove = FALSE) %>% 
  filter(state %in% c("washington", "oregon")) %>% 
  mutate(state = toupper(substr(state, 1, 2)))

counties$area <- as.numeric(st_area(counties))

# Wenatchee does not seem to have independent field office
pnw_field_offices <- field_offices %>% 
  filter(state != "AK",
         city != "Wenatchee") %>% 
  mutate(label = paste0(city, " (", site_code, ")"))

st_crs(pnw) <- 4326
st_crs(pnw$geom) <- 4326
# st_crs(sites) <- 4326
# st_crs(sites$geometry) <- 4326

m1 <- ggplot(data = pnw) +
  geom_sf() +
  geom_sf(data = pnw, fill = NA) + 
  geom_sf(data = counties) + 
  geom_sf(data = pnw_field_offices, color = "black") +
  geom_label_repel(data = pnw_field_offices, aes(label = label, geometry = geometry),
                   segment.color = 'black',
                   min.segment.length = 0,
                   segment.size = .5,,
                   stat = "sf_coordinates",
                force = 5) +
  scale_fill_viridis_c() +
  pnw_coords +
  theme_minimal() +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        panel.grid = element_blank()) +
  # facet_wrap(~apprehension_year) +
  labs(title = 'ICE field offices in the Pacific Northwest',
       caption = "Wenatchee (WNT) field office located in Yakima")

m1

```

```{r map_wa_2025, echo=TRUE, warning=FALSE}

wa <- states %>% 
  filter(ID == "Washington")

counties <- st_as_sf(maps::map("county", plot = FALSE, fill = TRUE))

counties <- counties %>% 
  separate(col = 'ID', into = c("state", "county"), sep = ",", remove = FALSE) %>% 
  filter(state == "washington") %>% 
  mutate(state = toupper(substr(state, 1, 2)))

counties$area <- as.numeric(st_area(counties))

counties <- left_join(counties, annual_arrests_pc, by = c("county" = "county_lower", "state" = "state")) %>% 
  drop_na()

counties <- counties %>%
  filter(cy == 2025)

wa_coords <- coord_sf(xlim = c(-125, -116.5), ylim = c(45, 49.5), expand = FALSE, crs = 4326)

# Wenatchee does not seem to have independent field office
wa_field_offices <- field_offices %>% 
  filter(state == "WA",
         city != "Wenatchee")

st_crs(wa) <- 4326
st_crs(wa$geom) <- 4326
# st_crs(sites) <- 4326
# st_crs(sites$geometry) <- 4326

m1 <- ggplot(data = wa) +
  geom_sf() +
  geom_sf(data = wa, fill = NA) + 
  geom_sf(data = counties, aes(fill = arrests_pc)) + 
  geom_sf(data = wa_field_offices, color = "black") +
  geom_label_repel(data = wa_field_offices, aes(label = city, geometry = geometry),
                   segment.color = 'black',
                   min.segment.length = 0,
                   segment.size = .5,,
                   stat = "sf_coordinates",
                force = 5) +
  scale_fill_viridis_c(limits = c(0, max(annual_arrests_pc$arrests_pc))) +
  wa_coords +
  theme_minimal() +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        panel.grid = element_blank()) +
  # facet_wrap(~apprehension_year) +
  labs(title = 'ICE arrests per capita',
       subtitle = "Jan.- Dec. 2025",
       fill = "Arrests per\n100,000",
       caption = general_caption)

m1

ggsave("../output/wa-ice-pc-2025.png", dpi=300)

```

```{r map_or_2025, echo=TRUE, warning=FALSE}

or <- states %>% 
  filter(ID == "Oregon")

counties <- st_as_sf(maps::map("county", plot = FALSE, fill = TRUE))

counties <- counties %>% 
  separate(col = 'ID', into = c("state", "county"), sep = ",", remove = FALSE) %>% 
  filter(state == "oregon") %>% 
  mutate(state = toupper(substr(state, 1, 2)))

counties$area <- as.numeric(st_area(counties))

counties <- left_join(counties, annual_arrests_pc, by = c("county" = "county_lower", "state" = "state")) %>% 
  drop_na()

counties <- counties %>%
  filter(cy == 2025)

or_field_offices <- field_offices %>% 
  filter(state == "OR")

st_crs(or) <- 4326
st_crs(or$geom) <- 4326
# st_crs(sites) <- 4326
# st_crs(sites$geometry) <- 4326

m1 <- ggplot(data = or) +
  geom_sf() +
  geom_sf(data = or, fill = NA) + 
  geom_sf(data = counties, aes(fill = arrests_pc)) + 
  geom_sf(data = or_field_offices, color = "black") +
  geom_label_repel(data = or_field_offices, aes(label = city, geometry = geometry),
                   segment.color = 'black',
                   min.segment.length = 0,
                   segment.size = .5,,
                   stat = "sf_coordinates",
                force = 1,
                force_pull = .25,
                nudge_x = .5,
                nudge_y = .5) +
  scale_fill_viridis_c(limits = c(0, max(annual_arrests_pc$arrests_pc))) +
  or_coords +
  theme_minimal() +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        panel.grid = element_blank()) +
  # facet_wrap(~apprehension_year) +
  labs(title = 'ICE arrests per capita',
       subtitle = "Jan.- Dec. 2025",
       fill = "Arrests per\n100,000",
       caption = general_caption)

m1

ggsave("../output/or-ice-pc-2025.png", dpi=300)


m2 <- ggplot(data = or) +
  geom_sf() +
  geom_sf(data = or, fill = NA) + 
  geom_sf(data = counties, aes(fill = arrests_pc)) + 
  geom_sf(data = or_field_offices, color = "black") +
  geom_label_repel(data = or_field_offices, aes(label = city, geometry = geometry),
                   segment.color = 'black',
                   min.segment.length = 0,
                   segment.size = .5,,
                   stat = "sf_coordinates",
                force = 1,
                force_pull = .25,
                nudge_x = .5,
                nudge_y = .5) +
  scale_fill_viridis_c(limits = c(0, or_pc_scale)) +
  or_coords +
  theme_minimal() +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        panel.grid = element_blank()) +
  # facet_wrap(~apprehension_year) +
  labs(title = 'ICE arrests per capita',
       subtitle = "Jan.- Dec. 2025",
       fill = "Arrests per\n100,000",
       caption = general_caption)

m2

ggsave("../output/or-ice-pc-2025-independent-scale.png", dpi=300)

```


```{r map_wa_annual, echo=TRUE, warning=FALSE}

wa <- states %>% 
  filter(ID == "Washington")

counties <- st_as_sf(maps::map("county", plot = FALSE, fill = TRUE))

counties <- counties %>% 
  separate(col = 'ID', into = c("state", "county"), sep = ",", remove = FALSE) %>% 
  filter(state == "washington") %>% 
  mutate(state = toupper(substr(state, 1, 2)))

counties$area <- as.numeric(st_area(counties))

counties <- left_join(counties, annual_arrests_pc, by = c("county" = "county_lower", "state" = "state")) %>% 
  drop_na()

wa_coords <- coord_sf(xlim = c(-125, -116.5), ylim = c(45, 49.5), expand = FALSE, crs = 4326)

st_crs(wa) <- 4326
st_crs(wa$geom) <- 4326
# st_crs(sites) <- 4326
# st_crs(sites$geometry) <- 4326

m1 <- ggplot(data = wa) +
  geom_sf() +
  geom_sf(data = wa, fill = NA) + 
  geom_sf(data = counties, aes(fill = arrests_pc)) + 
  # geom_sf_label(data = counties_rank, aes(label = str_to_title(county), background = NA), force = 100) +
  scale_fill_viridis_c() +
  wa_coords +
  theme_minimal() +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        panel.grid = element_blank()) +
  facet_wrap(~cy) +
  labs(title = 'ICE arrests per capita',
       subtitle = "2022 - 2025",
       fill = "Arrests per\n100,000",
       caption = general_caption)

m1

```

```{r map_or, echo=TRUE, warning=FALSE}

or <- states %>% 
  filter(ID == "Oregon")

counties <- st_as_sf(maps::map("county", plot = FALSE, fill = TRUE))

counties <- counties %>% 
  separate(col = 'ID', into = c("state", "county"), sep = ",", remove = FALSE) %>% 
  filter(state == "oregon") %>% 
  mutate(state = toupper(substr(state, 1, 2)))

counties$area <- as.numeric(st_area(counties))

counties <- left_join(counties, annual_arrests_pc, by = c("county" = "county_lower", "state" = "state")) %>% 
  drop_na()

st_crs(or) <- 4326
st_crs(or$geom) <- 4326
# st_crs(sites) <- 4326
# st_crs(sites$geometry) <- 4326

m1 <- ggplot(data = or) +
  geom_sf() +
  geom_sf(data = or, fill = NA) + 
  geom_sf(data = counties, aes(fill = arrests_pc)) + 
  scale_fill_viridis_c() +
  or_coords +
  theme_minimal() +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        panel.grid = element_blank()) +
  facet_wrap(~cy) +
  labs(title = 'ICE arrests per capita',
       subtitle = "per I-213 forms",
       caption = general_caption)

m1

```


```{r map_wa_or, echo=TRUE, warning=FALSE}

pnw <- states %>% 
  filter(ID %in% c('Washington', 'Oregon'))

counties <- st_as_sf(maps::map("county", plot = FALSE, fill = TRUE))

counties <- counties %>% 
  separate(col = 'ID', into = c("state", "county"), sep = ",", remove = FALSE) %>% 
  filter(state %in% c('washington', 'oregon'))  %>% 
  mutate(state = toupper(substr(state, 1, 2)))

counties$area <- as.numeric(st_area(counties))

counties <- left_join(counties, annual_arrests_pc, by = c("county" = "county_lower", "state")) %>% 
  drop_na()

st_crs(pnw) <- 4326
st_crs(pnw$geom) <- 4326
# st_crs(sites) <- 4326
# st_crs(sites$geometry) <- 4326

m1 <- ggplot(data = pnw) +
  geom_sf() +
  geom_sf(data = pnw, fill = NA) + 
  geom_sf(data = counties, aes(fill = arrests_pc)) + 
  scale_fill_viridis_c() +
  pnw_coords +
  theme_minimal() +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        panel.grid = element_blank()) +
  facet_wrap(~cy, nrow=1) +
  labs(title = 'ICE arrests per capita',
       subtitle = "per I-213 forms",
       caption = general_caption)

m1

```

```{r map_wa_or_2025, echo=TRUE, warning=FALSE}

pnw <- states %>% 
  filter(ID %in% c('Washington', 'Oregon'))

counties <- st_as_sf(maps::map("county", plot = FALSE, fill = TRUE))

counties <- counties %>% 
  separate(col = 'ID', into = c("state", "county"), sep = ",", remove = FALSE) %>% 
  filter(state %in% c('washington', 'oregon'))  %>% 
  mutate(state = toupper(substr(state, 1, 2)))

counties$area <- as.numeric(st_area(counties))

counties <- left_join(counties, annual_arrests_pc, by = c("county" = "county_lower", "state")) %>% 
  drop_na() %>% 
  filter(cy == 2025)

st_crs(pnw) <- 4326
st_crs(pnw$geom) <- 4326
# st_crs(sites) <- 4326
# st_crs(sites$geometry) <- 4326

m1 <- ggplot(data = pnw) +
  geom_sf() +
  geom_sf(data = pnw, fill = NA) + 
  geom_sf(data = counties, aes(fill = arrests_pc)) + 
  #   geom_sf(data = or_field_offices, color = "black") +
  # geom_label_repel(data = field_offices, aes(label = city, geometry = geometry),
  #                  segment.color = 'black',
  #                  min.segment.length = 0,
  #                  segment.size = .5,,
  #                  stat = "sf_coordinates",
  #               force = 1,
  #               force_pull = .25,
  #               nudge_x = .5,
  #               nudge_y = .5) +
  scale_fill_viridis_c() +
  pnw_coords +
  theme_minimal() +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        panel.grid = element_blank()) +
  labs(title = 'ICE arrests per capita',
       subtitle = "Jan.-Dec. 2025",
       fill = "Arrests per\n100,000",
       caption = general_caption)

m1

# ggsave(file="../output/pnw_per_capita.png", width=6, height=8, dpi=300)

```


### Quarterly analysis 2025

```{r map_wa_or_quarterly_per_capita, echo=TRUE, warning=FALSE}

pnw <- states %>% 
  filter(ID %in% c('Washington', 'Oregon'))

counties <- st_as_sf(maps::map("county", plot = FALSE, fill = TRUE))

counties <- counties %>% 
  separate(col = 'ID', into = c("state", "county"), sep = ",", remove = FALSE) %>% 
  filter(state %in% c('washington', 'oregon'))  %>% 
  mutate(state = toupper(substr(state, 1, 2)))

counties$area <- as.numeric(st_area(counties))

counties <- left_join(counties, quarterly_arrests_pc, by = c("county" = "county_lower", "state")) %>% 
  drop_na()

st_crs(pnw) <- 4326
st_crs(pnw$geom) <- 4326
# st_crs(sites) <- 4326
# st_crs(sites$geometry) <- 4326

plot_county_pc <- function(data_to_plot) {ggplot(data = pnw) +
  geom_sf() +
  geom_sf(data = pnw, fill = NA) + 
  geom_sf(data = data_to_plot, aes(fill = arrests_pc)) + 
  scale_fill_viridis_c(limits = c(0, 75)) +
  pnw_coords +
  theme_minimal() +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        panel.grid = element_blank()) +
  facet_wrap(~cy_quarter, nrow=1) +
  labs(title = 'ICE arrests per capita',
       subtitle = "per I-213 forms",
       caption = "I-213 data through Dec. 2025\n2022-2024 population estimates via U.S. census\n2024 estimate used for 2025")}

p1 <- counties %>% 
  filter(cy_quarter == "2025.1") %>% 
  plot_county_pc()

ggsave("../output/2025-1-pc.png")

p2 <- counties %>% 
  filter(cy_quarter == "2025.2") %>% 
  plot_county_pc()

ggsave("../output/2025-2-pc.png")

p3 <- counties %>% 
  filter(cy_quarter == "2025.3") %>% 
  plot_county_pc()

ggsave("../output/2025-3-pc.png")

p4 <- counties %>% 
  filter(cy_quarter == "2025.4") %>% 
  plot_county_pc()

ggsave("../output/2025-4-pc.png")

counties %>% 
  filter(cy == 2025) %>% 
  plot_county_pc() +
  facet_wrap(~cy_quarter, nrow=1)

```

```{r map_wa_or_quarterly_total_arrests, echo=TRUE, warning=FALSE}

wa <- states %>% 
  filter(ID %in% c('Washington', 'Oregon'))

counties <- st_as_sf(maps::map("county", plot = FALSE, fill = TRUE))

counties <- counties %>% 
  separate(col = 'ID', into = c("state", "county"), sep = ",", remove = FALSE) %>% 
  filter(state %in% c('washington', 'oregon'))  %>% 
  mutate(state = toupper(substr(state, 1, 2)))

counties$area <- as.numeric(st_area(counties))

counties <- left_join(counties, quarterly_arrests_pc, by = c("county" = "county_lower", "state")) %>% 
  drop_na()

st_crs(wa) <- 4326
st_crs(wa$geom) <- 4326
# st_crs(sites) <- 4326
# st_crs(sites$geometry) <- 4326

plot_county_pc <- function(data_to_plot) {ggplot(data = wa) +
  geom_sf() +
  geom_sf(data = wa, fill = NA) + 
  geom_sf(data = data_to_plot, aes(fill = arrests)) + 
  scale_fill_viridis_c(limits = c(0, 650)) +
  pnw_coords +
  theme_minimal() +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        panel.grid = element_blank()) +
  facet_wrap(~cy_quarter, nrow=1) +
  labs(title = 'ICE arrests',
       subtitle = "per I-213 forms",
       caption = "I-213 data through Dec. 2025")}

counties %>% 
  filter(cy == 2025) %>% 
  plot_county_pc() +
  facet_wrap(~cy_quarter, nrow=1)

```

```{r map_or_quarterly, echo=TRUE, warning=FALSE}

or <- states %>% 
  filter(ID == "Oregon")

counties <- st_as_sf(maps::map("county", plot = FALSE, fill = TRUE))

counties <- counties %>% 
  separate(col = 'ID', into = c("state", "county"), sep = ",", remove = FALSE) %>% 
  filter(state == "oregon") %>% 
  mutate(state = toupper(substr(state, 1, 2)))

counties$area <- as.numeric(st_area(counties))

counties <- left_join(counties, quarterly_arrests_pc, by = c("county" = "county_lower", "state" = "state")) %>% 
  drop_na() %>% 
  filter(cy == 2025) %>% 
  mutate(cy_quarter = str_replace(cy_quarter, "\\.", "Q"),
         months = case_when(cy_quarter == "2025Q1" ~ "Jan. - Mar.",
                            cy_quarter == "2025Q2" ~ "Apr. - Jun.",
                            cy_quarter == "2025Q3" ~ "Jul. - Sep.",
                            cy_quarter == "2025Q4" ~ "Oct. - Dec."))

st_crs(or) <- 4326
st_crs(or$geom) <- 4326
# st_crs(sites) <- 4326
# st_crs(sites$geometry) <- 4326

m1 <- ggplot(data = or) +
  geom_sf() +
  geom_sf(data = or, fill = NA) + 
  geom_sf(data = counties, aes(fill = arrests_pc)) + 
  scale_fill_viridis_c() +
  or_coords +
  theme_minimal() +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        panel.grid = element_blank()) +
  facet_wrap(~months) +
  labs(title = 'Quarterly ICE arrests per capita, Oregon counties',
       subtitle = "Jan. - Dec. 2025",
       fill = "Arrests per\n100,000",
       caption = general_caption)

m1

ggsave("../output/or-quarterly-arrests-pc-2025.png")

```


# Discrete arrest sites

```{r}

counties <- counties %>%
  separate(col = 'ID', into = c("state", "county"), sep = ",", remove = FALSE) %>%
  filter(state %in% c('washington', 'oregon'))

counties$area <- as.numeric(st_area(counties))

dat <- df %>% 
  filter(state %in% c("WA", "OR"),
         apprehension_date >= "2022-01-01") %>% 
  select(long, lat, site_code, cy) %>% 
  na.omit()

sites <- st_as_sf(dat, coords = c("long", "lat"),
                   crs = 4326, agr = "constant") %>% 
  count(geometry, cy, site_code) %>% 
  arrange(cy, site_code, n)

# wa_coords <- coord_sf(xlim = c(-125, -116.5), ylim = c(41.5, 49.5), expand = FALSE, crs = 4326)
# 
# st_crs(dat) <- 4326
# st_crs(dat$geom) <- 4326
st_crs(sites) <- 4326
st_crs(sites$geometry) <- 4326

plot_sites <- function(data_to_plot) {ggplot(data = wa) +
  geom_sf() +
  geom_sf(data = wa, fill = NA) + 
  geom_sf(data = counties, fill = NA) + 
  geom_sf(data = data_to_plot, shape = 21, color = "black", aes( fill = site_code, size = n), alpha=.7) +
  # geom_sf(data = field_offices, color = "black") +
  #   geom_label_repel(data = field_offices, aes(label = city, geometry = geometry),
  #                  segment.color = 'black',
  #                  min.segment.length = 0,
  #                  segment.size = .5,,
  #                  stat = "sf_coordinates",
  #               force = 5) +
  pnw_coords +
  facet_wrap(~cy, nrow=1) +
  theme_minimal() +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        panel.grid = element_blank()) +
  scale_size_continuous(limits = c(1, 600)) +
  labs(title = 'ICE arrests per year by field office',
       subtitle = "per I-213 forms",
       caption = "I-213 data through Dec. 2025")}

p1 <- sites %>% 
  filter(cy == 2025) %>% 
  plot_sites()

p1

p2 <- sites %>% 
  plot_sites()

p2

```

```{r}

counties <- counties %>%
  separate(col = 'ID', into = c("state", "county"), sep = ",", remove = FALSE) %>%
  filter(state %in% c('washington', 'oregon'))

counties$area <- as.numeric(st_area(counties))

sites <- df %>% 
  filter(state %in% c("WA", "OR"),
         apprehension_date >= "2025-01-01") %>% 
  select(long, lat, site_code, cy_quarter) %>% 
  na.omit()

sites <- st_as_sf(sites, coords = c("long", "lat"),
                   crs = 4326, agr = "constant") %>% 
  count(geometry, cy_quarter, site_code)

# wa_coords <- coord_sf(xlim = c(-125, -116.5), ylim = c(41.5, 49.5), expand = FALSE, crs = 4326)
# 
# st_crs(dat) <- 4326
# st_crs(dat$geom) <- 4326
st_crs(sites) <- 4326
st_crs(sites$geometry) <- 4326

plot_sites <- function(data_to_plot) {ggplot(data = wa) +
  geom_sf() +
  geom_sf(data = wa, fill = NA) + 
  geom_sf(data = counties, fill = NA) + 
  geom_sf(data = data_to_plot, shape = 21, color = "black", aes( fill = site_code, size = n), alpha=.7) +
  pnw_coords +
  facet_wrap(~cy_quarter, nrow=1) +
  theme_minimal() +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        panel.grid = element_blank()) +
  scale_size_continuous(limits = c(1, 600)) +
  labs(title = 'ICE arrests per quarter',
       subtitle = "per I-213 forms",
       caption = "I-213 data through Dec. 2025")}

p1 <- sites %>%
  filter(cy_quarter == "2025.1") %>% 
  plot_sites

ggsave("../output/2025-1.png")

p2 <- sites %>%
  filter(cy_quarter == "2025.2") %>% 
  plot_sites

ggsave("../output/2025-2.png")

p3 <- sites %>%
  filter(cy_quarter == "2025.3") %>% 
  plot_sites

ggsave("../output/2025-3.png")

p4 <- sites %>%
  filter(cy_quarter == "2025.4") %>% 
  plot_sites

ggsave("../output/2025-4.png")

```

```{r}

counties <- counties %>%
  separate(col = 'ID', into = c("state", "county"), sep = ",", remove = FALSE) %>%
  filter(state %in% c('washington', 'oregon'))

counties$area <- as.numeric(st_area(counties))

sites <- df %>% 
  filter(state %in% c("WA", "OR"),
         apprehension_date >= "2025-01-01") %>% 
  select(long, lat, cy_quarter) %>% 
  na.omit()

sites <- st_as_sf(sites, coords = c("long", "lat"),
                   crs = 4326, agr = "constant") %>% 
  count(geometry, cy_quarter) %>% 
  arrange(cy_quarter, n)

# wa_coords <- coord_sf(xlim = c(-125, -116.5), ylim = c(41.5, 49.5), expand = FALSE, crs = 4326)
# 
# st_crs(dat) <- 4326
# st_crs(dat$geom) <- 4326
st_crs(sites) <- 4326
st_crs(sites$geometry) <- 4326

plot_sites <- function(data_to_plot) {ggplot(data = wa) +
  geom_sf() +
  geom_sf(data = wa, fill = NA) + 
  geom_sf(data = counties, fill = NA) + 
  geom_sf(data = data_to_plot, aes(size = n), shape = 21, fill="red", color = "black",alpha=.7) +
  pnw_coords +
  facet_wrap(~cy_quarter, nrow=1) +
  theme_minimal() +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        panel.grid = element_blank()) +
  scale_size_continuous(limits = c(1, 600)) +
  labs(title = 'ICE arrests per quarter',
       subtitle = "per I-213 forms",
       caption = "I-213 data through Dec. 2025")}

# sites %>% plot_sites()

sites %>%
  # filter(cy_quarter == "2025.4") %>% 
  plot_sites() +
  facet_wrap(~cy_quarter, nrow=1)

```


```{r pdx_zoom_cy_quarter}

counties <- counties %>%
  separate(col = 'ID', into = c("state", "county"), sep = ",", remove = FALSE) %>%
  filter(state %in% c('washington', 'oregon'))

counties$area <- as.numeric(st_area(counties))

sites <- df %>% 
  filter(state %in% c("WA", "OR"),
         apprehension_date >= "2025-01-01",
         site_code == "POO") %>% 
  select(long, lat, cy_quarter) %>% 
  na.omit()

sites <- st_as_sf(sites, coords = c("long", "lat"),
                   crs = 4326, agr = "constant") %>% 
  count(geometry, cy_quarter) %>% 
  arrange(cy_quarter, n)


# 
# st_crs(dat) <- 4326
# st_crs(dat$geom) <- 4326
st_crs(sites) <- 4326
st_crs(sites$geometry) <- 4326

plot_sites <- function(data_to_plot) {ggplot(data = wa) +
  geom_sf() +
  geom_sf(data = wa, fill = NA, linewidth=.75) + 
  geom_sf(data = counties, fill = NA) + 
  geom_sf(data = data_to_plot, aes(size = n), shape = 21, color = "black", fill = "red", alpha=.7) +
  pdx_coords +
  facet_wrap(~cy_quarter, nrow=1) +
  theme_minimal() +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        panel.grid = element_blank()) +
  scale_size_continuous(limits = c(1, 600)) +
  labs(title = 'Operation "Black Rose"',
       subtitle = "per I-213 forms",
       caption = "I-213 data through Dec. 2025")}

p1 <- sites %>%
  # filter(cy_quarter == "2025.4") %>% 
  plot_sites +
  facet_wrap(~cy_quarter, nrow=2)

p1

```

```{r pdx_zoom_weekly}

counties <- counties %>%
  separate(col = 'ID', into = c("state", "county"), sep = ",", remove = FALSE) %>%
  filter(state %in% c('washington', 'oregon'))

counties$area <- as.numeric(st_area(counties))

sites <- df %>% 
  filter(state %in% c("WA", "OR"),
         apprehension_date >= "2025-10-06",
         apprehension_date <= "2025-12-28") %>% 
  select(long, lat, week) %>% 
  na.omit()

sites <- st_as_sf(sites, coords = c("long", "lat"),
                   crs = 4326, agr = "constant") %>% 
  count(geometry, week)


# 
# st_crs(dat) <- 4326
# st_crs(dat$geom) <- 4326
st_crs(sites) <- 4326
st_crs(sites$geometry) <- 4326

pdx_tight_coords <- coord_sf(xlim = c(-124, -121), ylim = c(44.5, 47), expand = FALSE, crs = 4326)

plot_sites <- function(data_to_plot) {ggplot(data = wa) +
  geom_sf() +
  geom_sf(data = wa, fill = NA) + 
  geom_sf(data = counties, fill = NA) + 
  geom_sf(data = data_to_plot, aes(size = n), shape = 21, color = "black", fill = "red", alpha=.7) +
  facet_wrap(~cy_quarter, nrow=1) +
    pdx_tight_coords +
  theme_minimal() +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        panel.grid = element_blank()) +
  scale_size_continuous(limits = c(1, 600)) +
  labs(title = 'Operation "Black Rose"',
       subtitle = "per I-213 forms",
       caption = "I-213 data through Dec. 2025")}

p1 <- sites %>%
  plot_sites +
  facet_wrap(~week)

p1

```

```{r king_county_zoom}

counties <- st_as_sf(maps::map("county", plot = FALSE, fill = TRUE))

counties <- counties %>%
  separate(col = 'ID', into = c("state", "county"), sep = ",", remove = FALSE) %>%
  filter(state %in% c('washington', 'oregon'))

counties$area <- as.numeric(st_area(counties))

sites <- df %>% 
  filter(state %in% c("WA"),
         apprehension_date >= "2025-01-01") %>% 
  select(long, lat) %>% 
  na.omit()

sites <- st_as_sf(sites, coords = c("long", "lat"),
                   crs = 4326, agr = "constant") %>% 
  count(geometry)


# 
# st_crs(dat) <- 4326
# st_crs(dat$geom) <- 4326
st_crs(sites) <- 4326
st_crs(sites$geometry) <- 4326

king_co_coords <- coord_sf(xlim = c(-124, -121), ylim = c(47, 48.5), expand = FALSE, crs = 4326)

plot_sites <- function(data_to_plot) {ggplot(data = wa) +
  geom_sf() +
  geom_sf(data = wa, fill = NA) + 
  geom_sf(data = counties, fill = NA) + 
  geom_sf(data = data_to_plot, aes(size = n), shape = 21, color = "black", fill = "red", alpha=.7) +
  # facet_wrap(~cy_quarter, nrow=1) +
    king_co_coords +
  theme_minimal() +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        panel.grid = element_blank()) +
  scale_size_continuous(limits = c(1, 600)) +
  labs(title = 'King County',
       subtitle = "per I-213 forms",
       caption = "I-213 data through Dec. 2025")}

p1 <- sites %>%
  plot_sites

p1

```



```{r}

p1 <- df %>% 
  filter(apprehension_date >= "2025-06-01",
         site_code == "POO") %>% 
  count(week) %>% 
  ggplot(aes(x = week, y = n)) +
  geom_line(color="red")

p1

p2 <- df %>% 
  filter(apprehension_date >= "2025-01-01",
         site_code == "POO") %>% 
  count(apprehension_date) %>% 
  ggplot(aes(x = apprehension_date, y = n)) +
  geom_col(color="red")

p2

```

# County by POO arrest

```{r map_wa_or, echo=TRUE, warning=FALSE}

wa <- states %>% 
  filter(ID %in% c('Washington', 'Oregon'))

counties <- st_as_sf(maps::map("county", plot = FALSE, fill = TRUE))

counties <- counties %>% 
  separate(col = 'ID', into = c("state", "county"), sep = ",", remove = FALSE) %>% 
  filter(state %in% c('washington', 'oregon'))  %>% 
  mutate(state = toupper(substr(state, 1, 2)))

counties$area <- as.numeric(st_area(counties))

dat <- df %>% 
  count(cy_quarter, state, county, site_code) %>% 
  filter(site_code == "POO") %>% 
  mutate(county_lower = tolower(str_replace(county, " County", "")))

counties <- left_join(counties, dat, by = c("county" = "county_lower", "state")) %>% 
  drop_na()

# counties <- counties %>%
#   filter(apprehension_year == 2025)

st_crs(wa) <- 4326
st_crs(wa$geom) <- 4326
# st_crs(sites) <- 4326
# st_crs(sites$geometry) <- 4326

plot_county_pc <- function(data_to_plot) {ggplot(data = wa) +
  geom_sf() +
  geom_sf(data = wa, fill = NA) + 
  geom_sf(data = data_to_plot, aes(fill = n)) + 
  scale_fill_viridis_c(limits = c(0, 600)) +
  pnw_coords +
  theme_minimal() +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        panel.grid = element_blank()) +
  labs(title = 'Total arrests by Portland ICE sub-office',
       subtitle = "per I-213 forms",
       caption = "I-213 data through Dec. 2025\n2022-2024 population estimates via U.S. census\n2024 estimate used for 2025")}

p1 <- counties %>% 
  filter(cy_quarter == "2025.1") %>% 
  plot_county_pc()

ggsave("../output/2025-1-poo.png")

p2 <- counties %>% 
  filter(cy_quarter == "2025.2") %>% 
  plot_county_pc()

ggsave("../output/2025-2-poo.png")

p3 <- counties %>% 
  filter(cy_quarter == "2025.3") %>% 
  plot_county_pc()

ggsave("../output/2025-3-poo.png")

p4 <- counties %>% 
  filter(cy_quarter == "2025.4") %>% 
  plot_county_pc()

ggsave("../output/2025-4-poo.png")

```